---
title: "Assignment 06.1"
subtitle: "Query a Database with SQL"
author: "Sai Karthikeyan, Sura"
date: "02/18/2025"
output: html_document
---

```{r setup, echo=FALSE}
# Load required libraries
knitr::opts_chunk$set(echo = TRUE)
library(DBI)
library(RSQLite)

# Connect to the SQLite database
dbCon <- dbConnect(RSQLite::SQLite(), "OrdersDB.sqlitedb.db")
```

## 1. What are the names and contacts of all suppliers, sorted by contact name?
```{sql suppliersContact, connection = dbCon}
SELECT SupplierName, ContactName, Phone
FROM Suppliers
ORDER BY ContactName;
```

## 2. What are the number of suppliers for each country? List the country and the number of suppliers renamed to 'NumSuppliers'.
```{sql suppliersPerCountry, connection = dbCon}
SELECT Country, COUNT(*) AS NumSuppliers
FROM Suppliers
GROUP BY Country
ORDER BY Country;
```

## 3. List the number of orders by customer ID and name, ordered from least to most, restricted to those customers who have placed at least one order.
```{sql ordersPerCustomer, connection = dbCon}
SELECT C.CustomerID, C.CustomerName, COUNT(O.OrderID) AS OrderCount
FROM Customers C
JOIN Orders O ON C.CustomerID = O.CustomerID
GROUP BY C.CustomerID
HAVING OrderCount > 0
ORDER BY OrderCount ASC;
```

## 4. What is the order id of the most recent order?
```{sql mostRecentOrderID, connection = dbCon}
SELECT OrderID
FROM Orders
ORDER BY OrderDate DESC
LIMIT 1;
```

## 5. What are the total number of unique orders delivered by each shipper, ordered alphabetically by shipper name?
```{sql ordersByShipper, connection = dbCon}
SELECT S.ShipperName, COUNT(DISTINCT O.OrderID) AS TotalOrders
FROM Shippers S
JOIN Orders O ON S.ShipperID = O.ShipperID
GROUP BY S.ShipperName
ORDER BY S.ShipperName;
```

## 6. How many employees never worked with a single customer, i.e., they appear in the Employees table but not in the Orders table?
```{sql employeesWithoutOrders, connection = dbCon}
SELECT COUNT(*) AS EmployeesWithoutOrders
FROM Employees E
LEFT JOIN Orders O ON E.EmployeeID = O.EmployeeID
WHERE O.OrderID IS NULL;
```

## 7. How many products contain "Louisiana" at the beginning of the name of the product?
```{sql louisianaProducts, connection = dbCon}
SELECT COUNT(*) AS LouisianaProducts
FROM Products
WHERE ProductName LIKE 'Louisiana%';
```

## 8. What is the total number of distinct countries to which at least one order was shipped?
```{sql countriesWithOrders, connection = dbCon}
SELECT COUNT(DISTINCT C.Country) AS DistinctCountryWithOrders
FROM Customers C
JOIN ORDERS O ON C.CustomerID = O.CustomerID;
```

## 9. What is the total amount (in terms of revenue) as well as the total number of orders sold by each employee? List the employee name, the total amount sold, and the total number of orders.
```{sql employeeSales, connection = dbCon}
SELECT E.FirstName, E.LastName, SUM(OD.Quantity * P.Price) AS TotalRevenue, COUNT(DISTINCT O.OrderID) AS TotalOrders
FROM Employees E
JOIN Orders O ON E.EmployeeID = O.EmployeeID
JOIN OrderDetails OD ON O.OrderID = OD.OrderID
JOIN Products P ON OD.ProductID = P.ProductID
GROUP BY E.EmployeeID
ORDER BY TotalRevenue DESC;
```

## 10. Which supplier sells the least number of products and supplies at least one product?
```{sql supplierWithLeastProducts, connection = dbCon}
WITH SupplierProductCounts AS (
    SELECT S.SupplierName, COUNT(P.ProductID) AS ProductCount
    FROM Suppliers S
    JOIN Products P ON S.SupplierID = P.SupplierID
    GROUP BY S.SupplierName
)
SELECT SupplierName
FROM SupplierProductCounts
WHERE ProductCount = (SELECT MIN(ProductCount) FROM SupplierProductCounts WHERE ProductCount > 0);
```

## 11. Which product was ordered most often by different customers?
```{sql mostOrderedProduct, connection = dbCon}
WITH ProductOrders AS (
    SELECT P.ProductName, COUNT(DISTINCT O.CustomerID) AS OrderCount
    FROM Products P
    JOIN OrderDetails OD ON P.ProductID = OD.ProductID
    JOIN Orders O ON OD.OrderID = O.OrderID
    GROUP BY P.ProductName
)
SELECT ProductName
FROM ProductOrders
WHERE OrderCount = (SELECT MAX(OrderCount) FROM ProductOrders);
```

## 12. Which product has the most number of units sold?
```{sql productMostUnitsSold, connection = dbCon}
WITH ProductSales AS (
    SELECT P.ProductName, SUM(OD.Quantity) AS TotalUnitsSold
    FROM Products P
    JOIN OrderDetails OD ON P.ProductID = OD.ProductID
    GROUP BY P.ProductName
)
SELECT ProductName
FROM ProductSales
WHERE TotalUnitsSold = (SELECT MAX(TotalUnitsSold) FROM ProductSales);
```

## 13. What is the total amount spent by all customers who live in either USA or Canada?
```{sql spendingInUSACanada, connection = dbCon}
SELECT SUM(P.Price * OD.Quantity) AS TotalAmountSpent
FROM Orders O
JOIN Customers C ON O.CustomerID = C.CustomerID
JOIN OrderDetails OD ON O.OrderID = OD.OrderID
JOIN Products P ON OD.ProductID = P.ProductID
WHERE C.Country IN ('USA', 'Canada');
```

## 14. What is the difference in spending between the country to which the most was sold versus the country to which the least was sold?
```{sql spendingDifference, connection = dbCon}
SELECT MAX(TotalSales) - MIN(TotalSales) AS SalesDifference
FROM(
  SELECT C.Country, SUM(P.Price * OD.Quantity) AS TotalSales
  FROM Customers C
  join Orders O on C.CustomerID = O.CustomerID
  join OrderDetails OD on O.OrderID = OD.OrderID
  join Products P on OD.ProductID = P.ProductID
  group by C.Country
);
```

## 15. Which country has the most customers?
```{sql countryMostCustomers, connection = dbCon}
SELECT Country, COUNT(CustomerID) AS NumCustomers
FROM Customers
GROUP BY Country
ORDER BY NumCustomers DESC
LIMIT 1;
```

## 16. Which employee sold the most in terms of "dollar amount" sold?
```{sql topEmployeeSales, connection = dbCon}
SELECT E.FirstName, E.LastName, SUM(OD.Quantity * P.Price) AS TotalRevenue
FROM Employees E
JOIN Orders O ON E.EmployeeID = O.EmployeeID
JOIN OrderDetails OD ON O.OrderID = OD.OrderID
JOIN Products P ON OD.ProductID = P.ProductID
GROUP BY E.EmployeeID
ORDER BY TotalRevenue DESC
LIMIT 1;
```

## 17. Which customers (name and ID) have never bought anything?
```{sql customersNeverBought, connection = dbCon}
SELECT CustomerID, CustomerName
FROM Customers
WHERE CustomerID NOT IN (SELECT DISTINCT CustomerID FROM Orders);
```

## 18. What is the average price of the products sold by each supplier?
```{sql avgPriceBySupplier, connection = dbCon}
SELECT S.SupplierName, AVG(P.Price) AS AverageProductPrice
FROM Suppliers S
JOIN Products P ON S.SupplierID = P.SupplierID
GROUP BY S.SupplierName;
```

## 19. To which customer was the most recent order sent?
```{sql mostRecentOrderCustomer, connection = dbCon}
SELECT C.CustomerName
FROM Orders O
JOIN Customers C ON O.CustomerID = C.CustomerID
ORDER BY O.OrderDate DESC
LIMIT 1;
```

## 20. From how many different (unique) suppliers do the products come from?
```{sql uniqueSuppliersCount, connection = dbCon}
SELECT COUNT(DISTINCT SupplierID) AS NumberOfSuppliers
FROM Products;
```

```{r echo=FALSE}
# close the database connection
dbDisconnect(dbCon)
```
